---
title: "mycodebookpractice"
output: html_document
date: "2025-09-30"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, load packages}
library(tidyverse)
library(summarytools)
library(survey)
```

```{r, load data}
load("~/code/NSDUH_2023.Rdata")
library(readr)
```

```{r, renaming}
 d <-"~/code/NSDUH_2023.Rdata"

 vars <-as_tibble(colnames(d))
```

#variables of interest, WHAT AGES DO RANGES ACTUALLY CORRELATE TO?
#demographics/confounders: QUESTID2, AGE3, CATAG2 (recoded 12-17=1), IRSEX, NEWRACE2, EDUHIGHCAT, HEALTH2 ??, IRFAMIN3 OR POVERTY3 ??,
#soft drugs: CIGFLAG, CGRFLAG, PIPFLAG, SMKLSSFLAG, NICVAPFLAG, ALCFLAG, CBDHMPFLAG, MRJFLAG, 0=NEVER USED, 1=EVER USED
#hard drugs: COCFLAG, CRKFLAG, HERFLAG, HALLUCFLAG, INHALFLAG, METHAMFLAG, PNRANYFLAG, TRQANYFLAG, STMANYFLAG, SEDANYFLAG, 0=NEVER USED, 1=EVER USED
#youth experiences aka suicidal thoughts in past year: YUSUITHKYR, .=18+/UNK, 1=YES, 2=NO, 3=IDK, 4=REFUSED
#adolescent depression pg504: lifetime mde YMDELT, .=+18/UNK, 1=YES, 2=NO; past year mde and alc use disorder YMDEAUD5YR, past year mde and illicit drug use disorder YMIUD5YANY, past year mde and substance use disorder YMSUD5YANY, .=+18/UNK, 0=NO, 1=YES


```{r, new data with vars of interest}
 s <- data %>%
   select(c(QUESTID2, IRCIGAGE, IRNICVAPAGE, IRSMKLSSTRY, IRCGRAGE, IRALCAGE, IRMJAGE, IRCOCAGE, IRHERAGE, IRMETHAMAGE, IRHALLUCAGE))
```

```{r, freq}
 freq(s$CIGTRY)
```

```{r, mean}
mean(s$CIGTRY)
```

```{r, median}
median(s$CIGTRY)
```

```{r, glimpse dataset}
glimpse(s)
```

```{r, cigtry-safe-recode}
#i dont think i like this...

# Safely coerce CIGTRY to numeric and keep only ages 1..65, then compute median and show counts
library(dplyr)

# prefer `s` if it exists (your analytic subset), otherwise use `d`
df_for_cig <- if (exists('s')) s else d

df_for_cig <- df_for_cig %>%
  mutate(CIGTRY_num = suppressWarnings(as.numeric(as.character(CIGTRY)))) %>%
  mutate(CIGTRY_clean = ifelse(CIGTRY_num >= 1 & CIGTRY_num <= 65, CIGTRY_num, NA_real_))

summary_tbl <- df_for_cig %>% summarise(
  total = n(),
  valid_1_65 = sum(!is.na(CIGTRY_clean)),
  median_1_65 = median(CIGTRY_clean, na.rm = TRUE)
)

print(summary_tbl)

cat('\nDistinct original CIGTRY codes and counts (top 50):\n')
print(head(sort(table(df_for_cig$CIGTRY_num), decreasing = TRUE), 50))

```

```{r, recode 991 to missing}

df_recoded <- df %>%
  mutate(IRCIGAGE = if_else(IRCIGAGE == 991, NA_integer_, IRCIGAGE))

print(df_recoded)
#>   ID IRCIGAGE
#> 1  1       15
#> 2  2       17
#> 3  3       NA
#> 4  4       14
#> 5  5       NA
#> 6  6       18

```

```{r, recode 991 to missing}
#attempt on IRCIGAGE, success, 

s_recoded <- s %>%
  mutate(IRCIGAGE = if_else(IRCIGAGE == 991, NA_integer_, IRCIGAGE))

print(s_recoded)

```

```{r, recode 991 to missing}
#attempt on all vars, 
#You can give across() a vector of the exact variable names you want to change. The ~ (tilde) creates a mini-function where .x stands in for each variable as it's being processed.
# List of variables to recode

vars_to_recode <- c("IRCIGAGE", "IRNICVAPAGE", "IRSMKLSSTRY", "IRCGRAGE", 
                    "IRALCAGE", "IRMJAGE", "IRCOCAGE", "IRHERAGE", 
                    "IRMETHAMAGE", "IRHALLUCAGE")

# Use across() to apply the same rule to all variables in the list
# The 'all_of()' function helps select columns from a character vector

s_recoded <- s %>%
  mutate(across(all_of(vars_to_recode), ~ if_else(.x == 991, NA_integer_, .x)))

print(s_recoded)

```

```{r, mean of s_recoded var}

#To calculate the average of a variable while ignoring the missing (NA) values, you use the argument na.rm = TRUE inside the mean() function.

# Calculate the mean of IRCIGAGE, removing NA values
summary_stats <- s_recoded %>%
  summarise(avg_cig_age = mean(IRCIGAGE, na.rm = TRUE))

print(summary_stats)

```
```{r, mean of all s_recoded vars}

means_of_all <- c("IRCIGAGE", "IRNICVAPAGE", "IRSMKLSSTRY", "IRCGRAGE", 
                    "IRALCAGE", "IRMJAGE", "IRCOCAGE", "IRHERAGE", 
                    "IRMETHAMAGE", "IRHALLUCAGE")

summary_stats <- s_recoded %>%
  summarise(across(all_of(means_of_all), ~ if_else(.x == 991, NA_integer_, .x)))
mean(IRCIGAGE, na.rm = TRUE

print(summary_stats)



