---
title: "wedv2"
output: html_document
date: "2025-11-19"
---
---
title: "Youth and Drug Use (Weighted Analysis)"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

# Load required packages
library(tidyverse)
library(survey)
library(summarytools)
library(gtsummary)
library(dplyr)
library(tableone)
library(table1)


# Load the dataset (ensure NSDUH_2023.Rdata is in your working directory)
load("NSDUH_2023.Rdata")

# Create a copy of the dataset
d <- data
```

#data prep and cleaning

```{r data_preparation}

# Select relevant variables
s <- d %>%
  select(
    # Core IDs & Weights
    QUESTID2, VEREP, VESTR_C, ANALWT2_C,
    
    # Demographics / Confounders
    AGE3, CATAG2, IRSEX, NEWRACE2, EDUSCHGRD2, POVERTY3,
    
    # Drug Variables (Exposures) (all coded as 0=never 1=ever)
    CIGFLAG, CGRFLAG, PIPFLAG, SMKLSSFLAG, NICVAPFLAG, ALCFLAG, CBDHMPFLAG, 
    MRJFLAG, COCFLAG, CRKFLAG, HERFLAG, HALLUCFLAG, INHALFLAG, METHAMFLAG, 
    PNRANYFLAG, TRQANYFLAG, STMANYFLAG, SEDANYFLAG,
    
    # Outcome Variables
    YMDELT, YMDEAUD5YR, YMIUD5YANY, YMSUD5YANY, 
    
    # Other Covariates
    AVGGRADE, PARLMTSN, EDFAM18, IMOTHER, IFATHER, YEYARGUP, YEPPROUD,
    YEYFGTSW, YEPGDJOB, YESCHACT, YECOMACT
  )

# Define drug variable lists
drug_vars <- c("CIGFLAG", "CGRFLAG", "PIPFLAG", "SMKLSSFLAG", "NICVAPFLAG", 
               "ALCFLAG", "CBDHMPFLAG", "MRJFLAG", "COCFLAG", "CRKFLAG", 
               "HERFLAG", "HALLUCFLAG", "INHALFLAG", "METHAMFLAG", 
               "PNRANYFLAG", "TRQANYFLAG", "STMANYFLAG", "SEDANYFLAG")

soft_drug_vars <- c("CIGFLAG", "CGRFLAG", "PIPFLAG", "SMKLSSFLAG", 
                    "NICVAPFLAG", "ALCFLAG", "CBDHMPFLAG", "MRJFLAG")

hard_drug_vars <- c("COCFLAG", "CRKFLAG", "HERFLAG", "HALLUCFLAG", 
                    "INHALFLAG", "METHAMFLAG", "PNRANYFLAG", "TRQANYFLAG", 
                    "STMANYFLAG", "SEDANYFLAG")

# Create composite drug use variables
s_clean <- s %>%
  mutate(
    # Main Exposure: Any drug use ever
    any_drug_ever_used = case_when(
      if_any(all_of(drug_vars), ~ . == 1) ~ 1,
      if_all(all_of(drug_vars), ~ . == 0) ~ 0,
      TRUE ~ NA_real_
    ),
    
    # Secondary: Soft drugs
    any_soft_drug_ever_used = case_when(
      if_any(all_of(soft_drug_vars), ~ . == 1) ~ 1,
      if_all(all_of(soft_drug_vars), ~ . == 0) ~ 0,
      TRUE ~ NA_real_
    ),
    
    # Secondary: Hard drugs
    any_hard_drug_ever_used = case_when(
      if_any(all_of(hard_drug_vars), ~ . == 1) ~ 1,
      if_all(all_of(hard_drug_vars), ~ . == 0) ~ 0,
      TRUE ~ NA_real_
    )
  )

# Preview the composite variables
glimpse(s_clean %>% select(any_drug_ever_used, any_soft_drug_ever_used, any_hard_drug_ever_used))

# Define confounders
confounder_vars <- c("AGE3", "IRSEX", "NEWRACE2", "EDUSCHGRD2", "POVERTY3", "AVGGRADE", "PARLMTSN",
                     "EDFAM18", "IMOTHER", "IFATHER", "YEYARGUP", "YEPPROUD",
                      "YEYFGTSW", "YEPGDJOB", "YESCHACT", "YECOMACT")

# Prepare analytic dataset for youth
youth_analytic <- s_clean %>%
  # Filter for youth (ages 12-17)
  filter(AGE3 %in% c(1, 2, 3)) %>%
  
  # Recode variables
  mutate(
    # Recode outcome: Major Depressive Episode (MDE)
    YMDELT_factor = factor(YMDELT, 
                           levels = c(1, 2), 
                           labels = c("Yes", "No")),
    
    # Recode demographics
    AGE3 = factor(AGE3, 
                  levels = c(1, 2, 3), 
                  labels = c("12-13", "14-15", "16-17")),
    
    IRSEX = factor(IRSEX, 
                   levels = c(1, 2), 
                   labels = c("Male", "Female")),
    
    # Factor any, soft and hard ever used
    any_drug_ever_used_factor = factor(any_drug_ever_used, 
                                       levels = c(0, 1), 
                                       labels = c("Never Used", "Ever Used")),
    
    any_soft_drug_ever_used_factor = factor(any_soft_drug_ever_used, 
                                            levels = c(0, 1), 
                                            labels = c("Never Used", "Ever Used")),
    
    any_hard_drug_ever_used_factor = factor(any_hard_drug_ever_used, 
                                            levels = c(0, 1), 
                                            labels = c("Never Used", "Ever Used"))
  ) %>%
  
  # Exclude unknowns/don't knows from EDUSCHGRD2 (set to NA if 98 or 99, then filter out)
  mutate(
    EDUSCHGRD2 = ifelse(EDUSCHGRD2 %in% c(98, 99), NA, EDUSCHGRD2)
  ) %>%
  filter(!is.na(EDUSCHGRD2)) %>%  # Explicitly filter out NA (unknowns) for counts
  
  # Factor education, race and mde
  mutate(
    EDUSCHGRD2 = as.factor(EDUSCHGRD2),
    NEWRACE2   = as.factor(NEWRACE2)
  ) %>%
  
  mutate(
    depepisode = ifelse(YMDELT_factor == "Yes", 1, 0)
  ) %>%
  
  # List-wise deletion: remove NAs in outcome, exposure, and confounders
  drop_na(YMDELT_factor, any_drug_ever_used_factor, all_of(confounder_vars))

# Report dropped rows
original_rows <- nrow(s_clean %>% filter(AGE3 %in% c(1, 2, 3)))
final_rows <- nrow(youth_analytic)
cat("Dropped", original_rows - final_rows, "rows due to NAs (out of", original_rows, "youth rows).\n")

# Check sample sizes and distributions (unweighted) - unknowns excluded
cat("--- Unweighted Age Distribution ---\n")
print(freq(youth_analytic$AGE3))

cat("\n--- Unweighted Drug Use Distribution ---\n")
print(freq(youth_analytic$any_drug_ever_used_factor))

cat("\n--- Unweighted Depression Outcome Distribution ---\n")
print(freq(youth_analytic$YMDELT_factor))

```

#survey design

```{r survey_design}
# Create survey design object
svy_analytic <- svydesign(
  id      = ~VEREP,
  strata  = ~VESTR_C,
  weights = ~ANALWT2_C,
  data    = youth_analytic,
  nest    = TRUE
)

# Print survey design summary
print(svy_analytic)
```

# unweighted counts & percents

```{r table1}

# Define variable labels
var_labels <- list(
  AGE3 ~ "Age Group",
  IRSEX ~ "Sex",
  NEWRACE2 ~ "Race/Ethnicity",
  EDUSCHGRD2 ~ "Education",
  POVERTY3 ~ "Poverty Level",
  AVGGRADE ~ "Average Grades",
  PARLMTSN ~ "Parental Monitoring"
)

# Generate unweighted Table 1 
tbl1 <- youth_analytic %>%
  tbl_summary(
    include = c(AGE3, IRSEX, NEWRACE2, EDUSCHGRD2, POVERTY3, AVGGRADE, 
                PARLMTSN),
    label = var_labels,
    statistic = list(all_categorical() ~ "{n} ({p}%)")
  ) %>%
  add_overall(
    col_label = "**Overall (N = {N})**",
    statistic = list(all_categorical() ~ "{n} ({p}%)")
  )

# Display the table (prints directly in RMarkdown)
tbl1

# Export table to CSV
write.csv(tbl1$table_body, "table1_unweighted.csv", row.names = FALSE)

```

# SB method: Weighted counts, percents, CIs
```{r}

# Weighted counts
weighted_counts <- svytable(~AGE3, svy_analytic)
weighted_counts
# Weighted proportions & 95% CI:
svyprop <- svymean(~factor(AGE3), design = svy_analytic)
svyprop
confint(svyprop)

# Combine results into a table
tbl <- data.frame(
  AGE3 = names(weighted_counts),
  Weighted_Count = as.numeric(weighted_counts),
  Proportion = coef(svyprop),
  SE = SE(svyprop),
  CI_Lower = confint(svyprop)[, 1],
  CI_Upper = confint(svyprop)[, 2]
)

tbl

########################################################################################################

# Weighted counts
weighted_counts <- svytable(~IRSEX, svy_analytic)
weighted_counts
# Weighted proportions & 95% CI:
svyprop <- svymean(~factor(IRSEX), design = svy_analytic)
svyprop
confint(svyprop)


# Combine results into a table
tbl <- data.frame(
  IRSEX = names(weighted_counts),
  Weighted_Count = as.numeric(weighted_counts),
  Proportion = coef(svyprop),
  CI_Lower = confint(svyprop)[, 1],
  CI_Upper = confint(svyprop)[, 2]
)

tbl

########################################################################################################

# Weighted counts
weighted_counts <- svytable(~EDUSCHGRD2, svy_analytic)
weighted_counts
# Weighted proportions & 95% CI:
svyprop <- svymean(~factor(EDUSCHGRD2), design = svy_analytic)
svyprop
confint(svyprop)


# Combine results into a table
tbl <- data.frame(
  EDUSCHGRD2 = names(weighted_counts),
  Weighted_Count = as.numeric(weighted_counts),
  Proportion = coef(svyprop),
  CI_Lower = confint(svyprop)[, 1],
  CI_Upper = confint(svyprop)[, 2]
)

tbl

########################################################################################################

# Weighted counts
weighted_counts <- svytable(~NEWRACE2, svy_analytic)
weighted_counts
# Weighted proportions & 95% CI:
svyprop <- svymean(~factor(NEWRACE2), design = svy_analytic)
svyprop
confint(svyprop)


# Combine results into a table
tbl <- data.frame(
  NEWRACE2 = names(weighted_counts),
  Weighted_Count = as.numeric(weighted_counts),
  Proportion = coef(svyprop),
  CI_Lower = confint(svyprop)[, 1],
  CI_Upper = confint(svyprop)[, 2]
)

tbl

########################################################################################################

# Weighted counts
weighted_counts <- svytable(~POVERTY3, svy_analytic)
weighted_counts
# Weighted proportions & 95% CI:
svyprop <- svymean(~factor(POVERTY3), design = svy_analytic)
svyprop
confint(svyprop)


# Combine results into a table
tbl <- data.frame(
  POVERTY3 = names(weighted_counts),
  Weighted_Count = as.numeric(weighted_counts),
  Proportion = coef(svyprop),
  CI_Lower = confint(svyprop)[, 1],
  CI_Upper = confint(svyprop)[, 2]
)

tbl

########################################################################################################

# Weighted counts
weighted_counts <- svytable(~AVGGRADE, svy_analytic)
weighted_counts
# Weighted proportions & 95% CI:
svyprop <- svymean(~factor(AVGGRADE), design = svy_analytic)
svyprop
confint(svyprop)


# Combine results into a table
tbl <- data.frame(
  AVGGRADE = names(weighted_counts),
  Weighted_Count = as.numeric(weighted_counts),
  Proportion = coef(svyprop),
  CI_Lower = confint(svyprop)[, 1],
  CI_Upper = confint(svyprop)[, 2]
)

tbl

########################################################################################################

# Weighted counts
weighted_counts <- svytable(~PARLMTSN, svy_analytic)
weighted_counts
# Weighted proportions & 95% CI:
svyprop <- svymean(~factor(PARLMTSN), design = svy_analytic)
svyprop
confint(svyprop)


# Combine results into a table
tbl <- data.frame(
  PARLMTSN = names(weighted_counts),
  Weighted_Count = as.numeric(weighted_counts),
  Proportion = coef(svyprop),
  CI_Lower = confint(svyprop)[, 1],
  CI_Upper = confint(svyprop)[, 2]
)

tbl

```

# Weighted counts, percents, CIs

```{r combined_weighted_table}
# List of variables to include in the combined table
variables <- c("AGE3", "IRSEX", "NEWRACE2", "EDUSCHGRD2", "POVERTY3", "AVGGRADE", "PARLMTSN")

# Function to compute weighted counts, proportions, and CIs for each variable
compute_weighted_table <- function(var, design) {
  # Weighted counts
  weighted_counts <- svytable(as.formula(paste("~", var)), design)
  
  # Weighted proportions
  svyprop <- svymean(as.formula(paste("~factor(", var, ")")), design = design)
  
  # Combine into data.frame
  data.frame(
    Variable = var,
    Category = names(weighted_counts),
    Weighted_Count = as.numeric(weighted_counts),
    Proportion = coef(svyprop),
    CI_Lower = confint(svyprop)[, 1],
    CI_Upper = confint(svyprop)[, 2]
  )
}

# Compute tables for all variables
all_tbls <- lapply(variables, compute_weighted_table, design = svy_analytic)

# Combine all tables into one
combined_tbl <- bind_rows(all_tbls)

# Display the combined table
combined_tbl

# Export to CSV
write.csv(combined_tbl, "weighted_table1.csv", row.names = FALSE)
```

# Confounder check - any drug

```{r confounder_tests}
# Define variables for testing
outcome_var <- "depepisode"
exposure_var <- "any_drug_ever_used_factor"
confounders_to_test <- c("AGE3", "IRSEX", "NEWRACE2", "EDUSCHGRD2", 
                         "POVERTY3", "AVGGRADE", "PARLMTSN",
                         "EDFAM18", "IMOTHER",    "IFATHER",    "YEYARGUP",
                         "YEPPROUD",   "YEYFGTSW",   "YEPGDJOB",  "YESCHACT", "YECOMACT" )



# Function to safely run svychisq and handle errors
safe_svychisq <- function(formula, design) {
  tryCatch({
    svychisq(formula, design = design)
  }, error = function(e) {
    message("Error in svychisq for ", deparse(formula), ": ", e$message)
    return(NULL)
  })
}

# Test confounders vs. outcome
cat("--- Confounder vs. Outcome (Depression) Tests ---\n")
conf_outcome_tests <- lapply(confounders_to_test, function(conf) {
  formula <- as.formula(paste("~", conf, "+", outcome_var))
  safe_svychisq(formula, svy_analytic)
})
names(conf_outcome_tests) <- confounders_to_test
print(conf_outcome_tests)

# Test confounders vs. exposure
cat("\n\n--- Confounder vs. Exposure (Drug Use) Tests ---\n")
conf_exposure_tests <- lapply(confounders_to_test, function(conf) {
  formula <- as.formula(paste("~", conf, "+", exposure_var))
  safe_svychisq(formula, svy_analytic)
})
names(conf_exposure_tests) <- confounders_to_test
print(conf_exposure_tests)
```

##testing new confounder code
```{r confounder_tests}
# Define variables for testing
outcome_var <- "depepisode"
exposure_var <- "any_drug_ever_used_factor"
confounders_to_test <- c("AGE3", "IRSEX", "NEWRACE2", "EDUSCHGRD2", 
                         "POVERTY3", "AVGGRADE", "PARLMTSN",
                         "EDFAM18", "IMOTHER", "IFATHER", "YEYARGUP",
                         "YEPPROUD", "YEYFGTSW", "YEPGDJOB", "YESCHACT", "YECOMACT")

# Function to safely run svychisq and handle errors
safe_svychisq <- function(formula, design) {
  tryCatch({
    svychisq(formula, design = design)
  }, error = function(e) {
    message("Error in svychisq for ", deparse(formula), ": ", e$message)
    return(NULL)
  })
}

# Function to extract results from svychisq objects
extract_chi_sq_results <- function(test_obj, confounder, level, type) {
  if (is.null(test_obj)) {
    return(NULL)
  }
  data.frame(
    Confounder = confounder,
    Level = as.character(level),  # Convert level to character to avoid type mismatch
    Type = type,
    Chi2 = test_obj$statistic,
    df = test_obj$parameter,
    p.value = test_obj$p.value
  )
}

# Initialize empty list for results
all_results <- list()

# First, add overall tests
cat("--- Overall Confounder vs. Outcome (Depression) Tests ---\n")
conf_outcome_tests <- lapply(confounders_to_test, function(conf) {
  formula <- as.formula(paste("~", conf, "+", outcome_var))
  safe_svychisq(formula, svy_analytic)
})
names(conf_outcome_tests) <- confounders_to_test

cat("\n\n--- Overall Confounder vs. Exposure (Drug Use) Tests ---\n")
conf_exposure_tests <- lapply(confounders_to_test, function(conf) {
  formula <- as.formula(paste("~", conf, "+", exposure_var))
  safe_svychisq(formula, svy_analytic)
})
names(conf_exposure_tests) <- confounders_to_test

# Extract overall results
for (conf in confounders_to_test) {
  outcome_test <- conf_outcome_tests[[conf]]
  exposure_test <- conf_exposure_tests[[conf]]
  
  if (!is.null(outcome_test)) {
    result <- extract_chi_sq_results(outcome_test, conf, "Overall", "vs Outcome (Depression)")
    all_results <- c(all_results, list(result))
  }
  
  if (!is.null(exposure_test)) {
    result <- extract_chi_sq_results(exposure_test, conf, "Overall", "vs Exposure (Drug Use)")
    all_results <- c(all_results, list(result))
  }
}

# Now, per-level tests for exposure vs outcome
for (conf in confounders_to_test) {
  # Get unique levels of the confounder in the data
  levels_conf <- unique(youth_analytic[[conf]])
  levels_conf <- levels_conf[!is.na(levels_conf)]  # Exclude NA if any
  
  # For each level of the confounder
  for (lvl in levels_conf) {
    # Subset the survey design to this level
    subset_design <- subset(svy_analytic, youth_analytic[[conf]] == lvl)
    
    # Check if subset has data
    if (nrow(subset_design$variables) == 0) next
    
    # Test exposure vs outcome in this subset
    formula <- as.formula(paste("~", exposure_var, "+", outcome_var))
    test_result <- safe_svychisq(formula, subset_design)
    
    # Extract results
    result <- extract_chi_sq_results(test_result, conf, lvl, "Exposure vs Outcome (within level)")
    if (!is.null(result)) {
      all_results <- c(all_results, list(result))
    }
  }
}

# Combine all results into one data.frame
all_chi_sq_results <- bind_rows(all_results)

# Display the combined chi-square table
all_chi_sq_results

# Export to CSV
write.csv(all_chi_sq_results, "chi_square_tests_complete.csv", row.names = FALSE)
```


# confounder check - soft drugs
```{r confounder_tests}
# Define variables for testing
outcome_var <- "depepisode"
exposure_var <- "any_soft_drug_ever_used_factor"
confounders_to_test <- c("AGE3", "IRSEX", "NEWRACE2", "EDUSCHGRD2", 
                         "POVERTY3", "AVGGRADE", "PARLMTSN",
                         "EDFAM18", "IMOTHER",    "IFATHER",    "YEYARGUP",
                         "YEPPROUD",   "YEYFGTSW",   "YEPGDJOB",  "YESCHACT", "YECOMACT" )



# Function to safely run svychisq and handle errors
safe_svychisq <- function(formula, design) {
  tryCatch({
    svychisq(formula, design = design)
  }, error = function(e) {
    message("Error in svychisq for ", deparse(formula), ": ", e$message)
    return(NULL)
  })
}

# Test confounders vs. outcome
cat("--- Confounder vs. Outcome (Depression) Tests ---\n")
conf_outcome_tests <- lapply(confounders_to_test, function(conf) {
  formula <- as.formula(paste("~", conf, "+", outcome_var))
  safe_svychisq(formula, svy_analytic)
})
names(conf_outcome_tests) <- confounders_to_test
print(conf_outcome_tests)

# Test confounders vs. exposure
cat("\n\n--- Confounder vs. Exposure (Drug Use) Tests ---\n")
conf_exposure_tests <- lapply(confounders_to_test, function(conf) {
  formula <- as.formula(paste("~", conf, "+", exposure_var))
  safe_svychisq(formula, svy_analytic)
})
names(conf_exposure_tests) <- confounders_to_test
print(conf_exposure_tests)
```

#confounder check - hard drugs

```{r confounder_tests}
# Define variables for testing
outcome_var <- "depepisode"
exposure_var <- "any_hard_drug_ever_used_factor"
confounders_to_test <- c("AGE3", "IRSEX", "NEWRACE2", "EDUSCHGRD2", 
                         "POVERTY3", "AVGGRADE", "PARLMTSN",
                         "EDFAM18", "IMOTHER",    "IFATHER",    "YEYARGUP",
                         "YEPPROUD",   "YEYFGTSW",   "YEPGDJOB",  "YESCHACT", "YECOMACT" )



# Function to safely run svychisq and handle errors
safe_svychisq <- function(formula, design) {
  tryCatch({
    svychisq(formula, design = design)
  }, error = function(e) {
    message("Error in svychisq for ", deparse(formula), ": ", e$message)
    return(NULL)
  })
}

# Test confounders vs. outcome
cat("--- Confounder vs. Outcome (Depression) Tests ---\n")
conf_outcome_tests <- lapply(confounders_to_test, function(conf) {
  formula <- as.formula(paste("~", conf, "+", outcome_var))
  safe_svychisq(formula, svy_analytic)
})
names(conf_outcome_tests) <- confounders_to_test
print(conf_outcome_tests)

# Test confounders vs. exposure
cat("\n\n--- Confounder vs. Exposure (Drug Use) Tests ---\n")
conf_exposure_tests <- lapply(confounders_to_test, function(conf) {
  formula <- as.formula(paste("~", conf, "+", exposure_var))
  safe_svychisq(formula, svy_analytic)
})
names(conf_exposure_tests) <- confounders_to_test
print(conf_exposure_tests)
```

# Regression - Aim 1 - SB version, ORs and CIs not printing

```{r}
# Check candidate confounder with OUTCOME
svychisq(~depepisode + AGE3, design = svy_analytic, statistic = "Chisq")
svychisq(~depepisode + IRSEX, design = svy_analytic, statistic = "Chisq")
svychisq(~depepisode + NEWRACE2, design = svy_analytic, statistic = "Chisq")
svychisq(~depepisode + EDUSCHGRD2, design = svy_analytic, statistic = "Chisq")
svychisq(~depepisode + POVERTY3, design = svy_analytic, statistic = "Chisq")
svychisq(~depepisode + AVGGRADE, design = svy_analytic, statistic = "Chisq")
svychisq(~depepisode + PARLMTSN, design = svy_analytic, statistic = "Chisq")
svychisq(~depepisode + EDFAM18, design = svy_analytic, statistic = "Chisq")
svychisq(~depepisode + IMOTHER, design = svy_analytic, statistic = "Chisq")
svychisq(~depepisode + IFATHER, design = svy_analytic, statistic = "Chisq")
svychisq(~depepisode + YEYARGUP, design = svy_analytic, statistic = "Chisq")
svychisq(~depepisode + YEPPROUD, design = svy_analytic, statistic = "Chisq")
svychisq(~depepisode + YEYFGTSW, design = svy_analytic, statistic = "Chisq")
svychisq(~depepisode + YEPGDJOB, design = svy_analytic, statistic = "Chisq")
svychisq(~depepisode + YESCHACT, design = svy_analytic, statistic = "Chisq")
svychisq(~depepisode + YECOMACT, design = svy_analytic, statistic = "Chisq")

# Check candidate confounders with EXPOSURE
##ANY drug ever used
svychisq(~any_drug_ever_used_factor + AGE3, design = svy_analytic, statistic = "Chisq")
svychisq(~any_drug_ever_used_factor + IRSEX, design = svy_analytic, statistic = "Chisq")
svychisq(~any_drug_ever_used_factor + NEWRACE2, design = svy_analytic, statistic = "Chisq")
svychisq(~any_drug_ever_used_factor + EDUSCHGRD2, design = svy_analytic, statistic = "Chisq")
svychisq(~any_drug_ever_used_factor + POVERTY3, design = svy_analytic, statistic = "Chisq")
svychisq(~any_drug_ever_used_factor + AVGGRADE, design = svy_analytic, statistic = "Chisq")
svychisq(~any_drug_ever_used_factor + PARLMTSN, design = svy_analytic, statistic = "Chisq")
svychisq(~any_drug_ever_used_factor + EDFAM18, design = svy_analytic, statistic = "Chisq")
svychisq(~any_drug_ever_used_factor + IMOTHER, design = svy_analytic, statistic = "Chisq")
svychisq(~any_drug_ever_used_factor + IFATHER, design = svy_analytic, statistic = "Chisq")
svychisq(~any_drug_ever_used_factor + YEYARGUP, design = svy_analytic, statistic = "Chisq")
svychisq(~any_drug_ever_used_factor + YEPPROUD, design = svy_analytic, statistic = "Chisq")
svychisq(~any_drug_ever_used_factor + YEYFGTSW, design = svy_analytic, statistic = "Chisq")
svychisq(~any_drug_ever_used_factor + YEPGDJOB, design = svy_analytic, statistic = "Chisq")
svychisq(~any_drug_ever_used_factor + YESCHACT, design = svy_analytic, statistic = "Chisq")
svychisq(~any_drug_ever_used_factor + YECOMACT, design = svy_analytic, statistic = "Chisq")

##SOFT drug ever used
svychisq(~any_soft_drug_ever_used_factor + AGE3, design = svy_analytic, statistic = "Chisq")
svychisq(~any_soft_drug_ever_used_factor + IRSEX, design = svy_analytic, statistic = "Chisq")
svychisq(~any_soft_drug_ever_used_factor + NEWRACE2, design = svy_analytic, statistic = "Chisq")
svychisq(~any_soft_drug_ever_used_factor + EDUSCHGRD2, design = svy_analytic, statistic = "Chisq")
svychisq(~any_soft_drug_ever_used_factor + POVERTY3, design = svy_analytic, statistic = "Chisq")
svychisq(~any_soft_drug_ever_used_factor + AVGGRADE, design = svy_analytic, statistic = "Chisq")
svychisq(~any_soft_drug_ever_used_factor + PARLMTSN, design = svy_analytic, statistic = "Chisq")
svychisq(~any_soft_drug_ever_used_factor + EDFAM18, design = svy_analytic, statistic = "Chisq")
svychisq(~any_soft_drug_ever_used_factor + IMOTHER, design = svy_analytic, statistic = "Chisq")
svychisq(~any_soft_drug_ever_used_factor + IFATHER, design = svy_analytic, statistic = "Chisq")
svychisq(~any_soft_drug_ever_used_factor + YEYARGUP, design = svy_analytic, statistic = "Chisq")
svychisq(~any_soft_drug_ever_used_factor + YEPPROUD, design = svy_analytic, statistic = "Chisq")
svychisq(~any_soft_drug_ever_used_factor + YEYFGTSW, design = svy_analytic, statistic = "Chisq")
svychisq(~any_soft_drug_ever_used_factor + YEPGDJOB, design = svy_analytic, statistic = "Chisq")
svychisq(~any_soft_drug_ever_used_factor + YESCHACT, design = svy_analytic, statistic = "Chisq")
svychisq(~any_soft_drug_ever_used_factor + YECOMACT, design = svy_analytic, statistic = "Chisq")

##HARD drug ever used
svychisq(~any_hard_drug_ever_used_factor + AGE3, design = svy_analytic, statistic = "Chisq")
svychisq(~any_hard_drug_ever_used_factor + IRSEX, design = svy_analytic, statistic = "Chisq")
svychisq(~any_hard_drug_ever_used_factor + NEWRACE2, design = svy_analytic, statistic = "Chisq")
svychisq(~any_hard_drug_ever_used_factor + EDUSCHGRD2, design = svy_analytic, statistic = "Chisq")
svychisq(~any_hard_drug_ever_used_factor + POVERTY3, design = svy_analytic, statistic = "Chisq")
svychisq(~any_hard_drug_ever_used_factor + AVGGRADE, design = svy_analytic, statistic = "Chisq")
svychisq(~any_hard_drug_ever_used_factor + PARLMTSN, design = svy_analytic, statistic = "Chisq")
svychisq(~any_hard_drug_ever_used_factor + EDFAM18, design = svy_analytic, statistic = "Chisq")
svychisq(~any_hard_drug_ever_used_factor + IMOTHER, design = svy_analytic, statistic = "Chisq")
svychisq(~any_hard_drug_ever_used_factor + IFATHER, design = svy_analytic, statistic = "Chisq")
svychisq(~any_hard_drug_ever_used_factor + YEYARGUP, design = svy_analytic, statistic = "Chisq")
svychisq(~any_hard_drug_ever_used_factor + YEPPROUD, design = svy_analytic, statistic = "Chisq")
svychisq(~any_hard_drug_ever_used_factor + YEYFGTSW, design = svy_analytic, statistic = "Chisq")
svychisq(~any_hard_drug_ever_used_factor + YEPGDJOB, design = svy_analytic, statistic = "Chisq")
svychisq(~any_hard_drug_ever_used_factor + YESCHACT, design = svy_analytic, statistic = "Chisq")
svychisq(~any_hard_drug_ever_used_factor + YECOMACT, design = svy_analytic, statistic = "Chisq")

# ^ where significant (p<0.05) in both OUTCOME AND EXPOSURE = Confounder
```


#regression, any drug every used
```{r}
###check to see which confounders stay and which go

########crude model adeu
# 1. Define the Adjusted Logistic Regression Model
m_adeucrude <- svyglm(depepisode ~ as.factor(any_drug_ever_used_factor), 
                      design = svy_analytic, 
                      family = quasibinomial)
# 2. Get the model summary (Log-Odds)
summary(m_adeucrude)
# 3. Extract the Log-Odds Estimates (Coefficients)
log_odds <- coef(m_adeucrude)
# 4. Extract the 95% Confidence Intervals for the Log-Odds
log_ci <- confint(m_adeucrude)
# 5. Combine the estimates and CIs into a single matrix
adeu_log <- cbind(
    Log_Odds = log_odds,
    Lower_CI = log_ci[, 1],
    Upper_CI = log_ci[, 2]
)
# 6. Exponentiate the entire matrix to get the Odds Ratios and their 95% CIs
adeu_OR_CI <- exp(adeu_log)
# 7. Print the final table of Odds Ratios and CIs
kable(adeu_OR_CI)

tbl_regression(m_adeucrude, exponentiate = TRUE) %>%
  bold_p()

#########biological model adeu
m_adeubiological <- svyglm(
    formula = depepisode ~ as.factor(any_drug_ever_used_factor) + AGE3 + IRSEX, 
    design = svy_analytic, 
    family = quasibinomial
)
summary(m_adeubiological)
log_odds <- coef(m_adeubiological)
log_ci <- confint(m_adeubiological)
adeubio_log <- cbind(
    Log_Odds = log_odds,
    Lower_CI = log_ci[, 1],
    Upper_CI = log_ci[, 2]
)
adeubio_OR_CI <- exp(adeubio_log)
kable(adeubio_OR_CI)

tbl_regression(m_adeubiological, exponentiate = TRUE) %>%
  bold_p()

#keep age and sex

#########social model adeu
m_adeusocial <- svyglm(depepisode ~ as.factor(any_drug_ever_used_factor) + # No comma here
                         AGE3 + 
                         IRSEX + 
                         NEWRACE2 + 
                         EDUSCHGRD2,
                      design = svy_analytic, 
                      family = quasibinomial,
                      na.action = na.omit         #added to exclude edu na's
                      )
summary(m_adeusocial)
log_odds <- coef(m_adeusocial)
log_ci <- confint(m_adeusocial)
adeusoc_log <- cbind(
    Log_Odds = log_odds,
    Lower_CI = log_ci[, 1],
    Upper_CI = log_ci[, 2]
)
adeusoc_OR_CI <- exp(adeusoc_log)
kable(adeusoc_OR_CI)

tbl_regression(m_adeusocial, exponentiate = TRUE) %>%
  bold_p()

#keep race and education
m_adeusocial <- svyglm(depepisode ~ as.factor(any_drug_ever_used_factor) + # No comma here
                         AGE3 + 
                         IRSEX + 
                         NEWRACE2 + 
                         EDUSCHGRD2,
                      design = svy_analytic, 
                      family = quasibinomial,
                      na.action = na.omit         #added to exclude edu na's
                      )
summary(m_adeusocial)
log_odds <- coef(m_adeusocial)
log_ci <- confint(m_adeusocial)
adeusoc_log <- cbind(
    Log_Odds = log_odds,
    Lower_CI = log_ci[, 1],
    Upper_CI = log_ci[, 2]
)
adeusoc_OR_CI <- exp(adeusoc_log)
kable(adeusoc_OR_CI)

tbl_regression(m_adeusocial, exponentiate = TRUE) %>%
  bold_p()

#########behavioral model adeu
m_adeubeh <- svyglm(depepisode ~ as.factor(any_drug_ever_used_factor) + # No comma here
                      AGE3 + 
                      IRSEX + 
                      NEWRACE2 + 
                      EDUSCHGRD2 + 
                      YEYARGUP + 
                      YEPPROUD +
                      YEYFGTSW + 
                      YEPGDJOB,
                      design = svy_analytic, 
                      family = quasibinomial,
                      na.action = na.omit         #added to exclude edu na's
                      )
summary(m_adeubeh)
log_odds <- coef(m_adeubeh)
log_ci <- confint(m_adeubeh)
adeubeh_log <- cbind(
    Log_Odds = log_odds,
    Lower_CI = log_ci[, 1],
    Upper_CI = log_ci[, 2]
)
adeubeh_OR_CI <- exp(adeubeh_log)
kable(adeubeh_OR_CI)

tbl_regression(m_adeubeh, exponentiate = TRUE) %>%
  bold_p()

#keep yeyargu
m_adeubeh <- svyglm(depepisode ~ as.factor(any_drug_ever_used_factor) + # No comma here
                      AGE3 + 
                      IRSEX + 
                      NEWRACE2 + 
                      EDUSCHGRD2 +
                      YEYARGUP,
                      design = svy_analytic, 
                      family = quasibinomial,
                      na.action = na.omit         #added to exclude edu na's
                      )
summary(m_adeubeh)
log_odds <- coef(m_adeubeh)
log_ci <- confint(m_adeubeh)
adeubeh_log <- cbind(
    Log_Odds = log_odds,
    Lower_CI = log_ci[, 1],
    Upper_CI = log_ci[, 2]
)
adeubeh_OR_CI <- exp(adeubeh_log)
kable(adeubeh_OR_CI)

tbl_regression(m_adeubeh, exponentiate = TRUE) %>%
  bold_p()

```

# regression, soft drug ever used
```{r}

########crude model sdeu
# 1. Define the Adjusted Logistic Regression Model
m_sdeucrude <- svyglm(depepisode ~ as.factor(any_soft_drug_ever_used_factor), 
                      design = svy_analytic, 
                      family = quasibinomial)
# 2. Get the model summary (Log-Odds)
summary(m_sdeucrude)
# 3. Extract the Log-Odds Estimates (Coefficients)
log_odds <- coef(m_sdeucrude)
# 4. Extract the 95% Confidence Intervals for the Log-Odds
log_ci <- confint(m_sdeucrude)
# 5. Combine the estimates and CIs into a single matrix
sdeu_log <- cbind(
    Log_Odds = log_odds,
    Lower_CI = log_ci[, 1],
    Upper_CI = log_ci[, 2]
)
# 6. Exponentiate the entire matrix to get the Odds Ratios and their 95% CIs
sdeu_OR_CI <- exp(sdeu_log)
# 7. Print the final table of Odds Ratios and CIs
kable(sdeu_OR_CI)

tbl_regression(m_sdeucrude, exponentiate = TRUE) %>%
  bold_p()

#########biological model sdeu
m_sdeubiological <- svyglm(
    formula = depepisode ~ as.factor(any_soft_drug_ever_used_factor) + 
              AGE3 + 
              IRSEX, 
          design = svy_analytic, 
          family = quasibinomial
          )
summary(m_sdeubiological)
log_odds <- coef(m_sdeubiological)
log_ci <- confint(m_sdeubiological)
sdeubio_log <- cbind(
    Log_Odds = log_odds,
    Lower_CI = log_ci[, 1],
    Upper_CI = log_ci[, 2]
)
sdeubio_OR_CI <- exp(sdeubio_log)
kable(sdeubio_OR_CI)

tbl_regression(m_sdeubiological, exponentiate = TRUE) %>%
  bold_p()

#########social model sdeu
m_sdeusocial <- svyglm(depepisode ~ as.factor(any_soft_drug_ever_used_factor) + 
                         AGE3 + 
                         IRSEX + 
                         NEWRACE2 + 
                         EDUSCHGRD2,
                      design = svy_analytic, 
                      family = quasibinomial,
                      na.action = na.omit         #added to exclude edu na's

                      )
summary(m_sdeusocial)
log_odds <- coef(m_sdeusocial)
log_ci <- confint(m_sdeusocial)
sdeusoc_log <- cbind(
    Log_Odds = log_odds,
    Lower_CI = log_ci[, 1],
    Upper_CI = log_ci[, 2]
)
sdeusoc_OR_CI <- exp(sdeusoc_log)
kable(sdeusoc_OR_CI)

tbl_regression(m_sdeusocial, exponentiate = TRUE) %>%
  bold_p()

#keep race and edu
m_sdeusocial <- svyglm(depepisode ~ as.factor(any_soft_drug_ever_used_factor) + 
                         AGE3 + 
                         IRSEX + 
                         NEWRACE2 + 
                         EDUSCHGRD2,
                      design = svy_analytic, 
                      family = quasibinomial,
                      na.action = na.omit         #added to exclude edu na's

                      )
summary(m_sdeusocial)
log_odds <- coef(m_sdeusocial)
log_ci <- confint(m_sdeusocial)
sdeusoc_log <- cbind(
    Log_Odds = log_odds,
    Lower_CI = log_ci[, 1],
    Upper_CI = log_ci[, 2]
)
sdeusoc_OR_CI <- exp(sdeusoc_log)
kable(sdeusoc_OR_CI)

tbl_regression(m_sdeusocial, exponentiate = TRUE) %>%
  bold_p()

#########behavioral model sdeu
m_sdeubeh <- svyglm(depepisode ~ as.factor(any_soft_drug_ever_used_factor) + 
                      AGE3 + 
                      IRSEX + 
                      NEWRACE2 + 
                      EDUSCHGRD2 + 
                      YEYARGUP + 
                      YEPPROUD +
                      YEYFGTSW + 
                      YEPGDJOB,
                      design = svy_analytic, 
                      family = quasibinomial,
                      na.action = na.omit         #added to exclude edu na's
                      )
summary(m_sdeubeh)
log_odds <- coef(m_sdeubeh)
log_ci <- confint(m_sdeubeh)
sdeubeh_log <- cbind(
    Log_Odds = log_odds,
    Lower_CI = log_ci[, 1],
    Upper_CI = log_ci[, 2]
)
sdeubeh_OR_CI <- exp(sdeubeh_log)
kable(sdeubeh_OR_CI)

tbl_regression(m_sdeubeh, exponentiate = TRUE) %>%
  bold_p()

#keep yeyar
m_sdeubeh <- svyglm(depepisode ~ as.factor(any_soft_drug_ever_used_factor) + 
                      AGE3 + 
                      IRSEX + 
                      NEWRACE2 + 
                      EDUSCHGRD2 + 
                      YEYARGUP,
                      design = svy_analytic, 
                      family = quasibinomial,
                      na.action = na.omit         #added to exclude edu na's
                      )
summary(m_sdeubeh)
log_odds <- coef(m_sdeubeh)
log_ci <- confint(m_sdeubeh)
sdeubeh_log <- cbind(
    Log_Odds = log_odds,
    Lower_CI = log_ci[, 1],
    Upper_CI = log_ci[, 2]
)
sdeubeh_OR_CI <- exp(sdeubeh_log)
kable(sdeubeh_OR_CI)

tbl_regression(m_sdeubeh, exponentiate = TRUE) %>%
  bold_p()
```

# regression, hard drug ever used
```{r}

########crude model hdeu
# 1. Define the Adjusted Logistic Regression Model
m_hdeucrude <- svyglm(depepisode ~ as.factor(any_hard_drug_ever_used_factor), 
                      design = svy_analytic, 
                      family = quasibinomial)
# 2. Get the model summary (Log-Odds)
summary(m_hdeucrude)
# 3. Extract the Log-Odds Estimates (Coefficients)
log_odds <- coef(m_hdeucrude)
# 4. Extract the 95% Confidence Intervals for the Log-Odds
log_ci <- confint(m_hdeucrude)
# 5. Combine the estimates and CIs into a single matrix
hdeu_log <- cbind(
    Log_Odds = log_odds,
    Lower_CI = log_ci[, 1],
    Upper_CI = log_ci[, 2]
)
# 6. Exponentiate the entire matrix to get the Odds Ratios and their 95% CIs
hdeu_OR_CI <- exp(hdeu_log)
# 7. Print the final table of Odds Ratios and CIs
kable(hdeu_OR_CI)

tbl_regression(m_hdeucrude, exponentiate = TRUE) %>%
  bold_p()

#########biological model hdeu
m_hdeubiological <- svyglm(
    formula = depepisode ~ as.factor(any_hard_drug_ever_used_factor) + 
              AGE3, 
          design = svy_analytic, 
          family = quasibinomial
          )
summary(m_hdeubiological)
log_odds <- coef(m_hdeubiological)
log_ci <- confint(m_hdeubiological)
hdeubio_log <- cbind(
    Log_Odds = log_odds,
    Lower_CI = log_ci[, 1],
    Upper_CI = log_ci[, 2]
)
hdeubio_OR_CI <- exp(hdeubio_log)
kable(hdeubio_OR_CI)

tbl_regression(m_hdeubiological, exponentiate = TRUE) %>%
  bold_p()

#keep age 

#########social model hdeu
m_hdeusocial <- svyglm(depepisode ~ as.factor(any_hard_drug_ever_used_factor) + 
                         AGE3 + 
                         NEWRACE2 + 
                         EDUSCHGRD2,
                      design = svy_analytic, 
                      family = quasibinomial,
                      na.action = na.omit         #added to exclude edu na's
                      )
summary(m_hdeusocial)
log_odds <- coef(m_hdeusocial)
log_ci <- confint(m_hdeusocial)
hdeusoc_log <- cbind(
    Log_Odds = log_odds,
    Lower_CI = log_ci[, 1],
    Upper_CI = log_ci[, 2]
)
hdeusoc_OR_CI <- exp(hdeusoc_log)
kable(hdeusoc_OR_CI)

tbl_regression(m_hdeusocial, exponentiate = TRUE) %>%
  bold_p()

#keep race/edu
m_hdeusocial <- svyglm(depepisode ~ as.factor(any_hard_drug_ever_used_factor) + 
                         AGE3 + 
                         NEWRACE2 + 
                         EDUSCHGRD2,
                      design = svy_analytic, 
                      family = quasibinomial,
                      na.action = na.omit         #added to exclude edu na's
                      )
summary(m_hdeusocial)
log_odds <- coef(m_hdeusocial)
log_ci <- confint(m_hdeusocial)
hdeusoc_log <- cbind(
    Log_Odds = log_odds,
    Lower_CI = log_ci[, 1],
    Upper_CI = log_ci[, 2]
)
hdeusoc_OR_CI <- exp(hdeusoc_log)
kable(hdeusoc_OR_CI)

tbl_regression(m_hdeusocial, exponentiate = TRUE) %>%
  bold_p()

#########behavioral model hdeu
m_hdeubeh <- svyglm(depepisode ~ as.factor(any_hard_drug_ever_used_factor) + 
                      AGE3 + 
                      NEWRACE2 + 
                      EDUSCHGRD2 + 
                      YEYARGUP + 
                      YEPPROUD +
                      YEYFGTSW + 
                      YEPGDJOB + 
                      YESCHACT,
                      design = svy_analytic, 
                      family = quasibinomial, 
                      na.action = na.omit         #added to exclude edu na's
                      )
summary(m_hdeubeh)
log_odds <- coef(m_hdeubeh)
log_ci <- confint(m_hdeubeh)
hdeubeh_log <- cbind(
    Log_Odds = log_odds,
    Lower_CI = log_ci[, 1],
    Upper_CI = log_ci[, 2]
)
hdeubeh_OR_CI <- exp(hdeubeh_log)
kable(hdeubeh_OR_CI)

tbl_regression(m_hdeubeh, exponentiate = TRUE) %>%
  bold_p()

#keep yeyar
m_hdeubeh <- svyglm(depepisode ~ as.factor(any_hard_drug_ever_used_factor) + 
                      AGE3 + 
                      NEWRACE2 + 
                      EDUSCHGRD2 + 
                      YEYARGUP,
                      design = svy_analytic, 
                      family = quasibinomial, 
                      na.action = na.omit         #added to exclude edu na's
                      )
summary(m_hdeubeh)
log_odds <- coef(m_hdeubeh)
log_ci <- confint(m_hdeubeh)
hdeubeh_log <- cbind(
    Log_Odds = log_odds,
    Lower_CI = log_ci[, 1],
    Upper_CI = log_ci[, 2]
)
hdeubeh_OR_CI <- exp(hdeubeh_log)
kable(hdeubeh_OR_CI)

tbl_regression(m_hdeubeh, exponentiate = TRUE) %>%
  bold_p()

```

# Overall Prevalence of outcome in analytic sample
```{r}

weighted_counts <- svytable(~ depepisode, svy_analytic)
svyprop <- svymean(~factor(depepisode), design = svy_analytic)
svyprop
confint(svyprop)

```

# Overall drug use behavior in analytic sample 
```{r}

# Overall 
weighted_counts <- svytable(~ any_drug_ever_used_factor, svy_analytic)
svyprop <- svymean(~factor(any_drug_ever_used_factor), design = svy_analytic)
svyprop
confint(svyprop)

#SOFT DRUGS
weighted_counts <- svytable(~ any_soft_drug_ever_used_factor, svy_analytic)
svyprop <- svymean(~factor(any_soft_drug_ever_used_factor), design = svy_analytic)
svyprop
confint(svyprop)

#HARD DRUGS
weighted_counts <- svytable(~ any_hard_drug_ever_used_factor, svy_analytic)
svyprop <- svymean(~factor(any_hard_drug_ever_used_factor), design = svy_analytic)
svyprop
confint(svyprop)

# Each drug (so we can see what drives this estimate - e.g., mainly alcohol and marijuana maybe)

#SOFT DRUGS
#ALCOHOL
weighted_counts <- svytable(~ ALCFLAG, svy_analytic)
svyprop <- svymean(~factor(ALCFLAG), design = svy_analytic)
svyprop
confint(svyprop)

#CIGS
weighted_counts <- svytable(~ CIGFLAG, svy_analytic)
svyprop <- svymean(~factor(CIGFLAG), design = svy_analytic)
svyprop
confint(svyprop)

#MARIJUANA
weighted_counts <- svytable(~ MRJFLAG, svy_analytic)
svyprop <- svymean(~factor(MRJFLAG), design = svy_analytic)
svyprop
confint(svyprop)

#CIGAR
weighted_counts <- svytable(~ CGRFLAG, svy_analytic)
svyprop <- svymean(~factor(CGRFLAG), design = svy_analytic)
svyprop
confint(svyprop)

#PIPE
weighted_counts <- svytable(~ PIPFLAG, svy_analytic)
svyprop <- svymean(~factor(PIPFLAG), design = svy_analytic)
svyprop
confint(svyprop)

#SMOKELESS
weighted_counts <- svytable(~ SMKLSSFLAG, svy_analytic)
svyprop <- svymean(~factor(SMKLSSFLAG), design = svy_analytic)
svyprop
confint(svyprop)

#NIC VAPES
weighted_counts <- svytable(~ NICVAPFLAG, svy_analytic)
svyprop <- svymean(~factor(NICVAPFLAG), design = svy_analytic)
svyprop
confint(svyprop)

#CBD/HEMP
weighted_counts <- svytable(~ CBDHMPFLAG, svy_analytic)
svyprop <- svymean(~factor(CBDHMPFLAG), design = svy_analytic)
svyprop
confint(svyprop)

#HARD DRUGS
#COCAINE
weighted_counts <- svytable(~ COCFLAG, svy_analytic)
svyprop <- svymean(~factor(COCFLAG), design = svy_analytic)
svyprop
confint(svyprop)

#CRACK
weighted_counts <- svytable(~ CRKFLAG, svy_analytic)
svyprop <- svymean(~factor(CRKFLAG), design = svy_analytic)
svyprop
confint(svyprop)

#HEROIN
weighted_counts <- svytable(~ HERFLAG, svy_analytic)
svyprop <- svymean(~factor(HERFLAG), design = svy_analytic)
svyprop
confint(svyprop)

#METH
weighted_counts <- svytable(~ METHAMFLAG, svy_analytic)
svyprop <- svymean(~factor(METHAMFLAG), design = svy_analytic)
svyprop
confint(svyprop)

#HALLUCINOGENS
weighted_counts <- svytable(~ HALLUCFLAG, svy_analytic)
svyprop <- svymean(~factor(HALLUCFLAG), design = svy_analytic)
svyprop
confint(svyprop)

#INHALANTS
weighted_counts <- svytable(~ INHALFLAG, svy_analytic)
svyprop <- svymean(~factor(INHALFLAG), design = svy_analytic)
svyprop
confint(svyprop)

#PAIN RELIEVERS
weighted_counts <- svytable(~ PNRANYFLAG, svy_analytic)
svyprop <- svymean(~factor(PNRANYFLAG), design = svy_analytic)
svyprop
confint(svyprop)

#TRANQS
weighted_counts <- svytable(~ TRQANYFLAG, svy_analytic)
svyprop <- svymean(~factor(TRQANYFLAG), design = svy_analytic)
svyprop
confint(svyprop)

#STIMULANTS
weighted_counts <- svytable(~ STMANYFLAG, svy_analytic)
svyprop <- svymean(~factor(STMANYFLAG), design = svy_analytic)
svyprop
confint(svyprop)

#SEDATIVES
weighted_counts <- svytable(~ SEDANYFLAG, svy_analytic)
svyprop <- svymean(~factor(SEDANYFLAG), design = svy_analytic)
svyprop
confint(svyprop)
```


#debugging
```{r debugging, eval=FALSE}
# Debugging: Check for variables with only one level (run if errors occur)
# Uncomment and run this chunk if tbl_svysummary or svychisq fails

cat("Checking categorical variables for single levels:\n")

check_vars <- c("AGE3", "IRSEX", "NEWRACE2", "EDUSCHGRD2", "POVERTY3", 
                "AVGGRADE", "PARLMTSN", "YMDELT_factor", "any_drug_ever_used_factor")

for (var in check_vars) {
  cat("\n---", var, "---\n")
  print(table(youth_analytic[[var]], useNA = "ifany"))
}
```
