---
title: "wedv2"
output: html_document
date: "2025-11-19"
---
---
title: "Youth and Drug Use (Weighted Analysis)"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

# Load required packages
library(tidyverse)
library(survey)
library(summarytools)
library(gtsummary)


# Load the dataset (ensure NSDUH_2023.Rdata is in your working directory)
load("NSDUH_2023.Rdata")

# Create a copy of the dataset
d <- data
```

```{r data_preparation}
# Select relevant variables
s <- d %>%
  select(
    # Core IDs & Weights
    QUESTID2, VEREP, VESTR_C, ANALWT2_C,
    
    # Demographics / Confounders
    AGE3, CATAG2, IRSEX, NEWRACE2, EDUSCHGRD2, POVERTY3,
    
    # Drug Variables (Exposures)
    CIGFLAG, CGRFLAG, PIPFLAG, SMKLSSFLAG, NICVAPFLAG, ALCFLAG, CBDHMPFLAG, 
    MRJFLAG, COCFLAG, CRKFLAG, HERFLAG, HALLUCFLAG, INHALFLAG, METHAMFLAG, 
    PNRANYFLAG, TRQANYFLAG, STMANYFLAG, SEDANYFLAG,
    
    # Outcome Variables
    YMDELT, YMDEAUD5YR, YMIUD5YANY, YMSUD5YANY, 
    
    # Other Covariates
    AVGGRADE, PARLMTSN, EDFAM18, IMOTHER, IFATHER, YEYARGUP, YEPPROUD,
    YEYFGTSW, YEPGDJOB, YESCHACT, YECOMACT
  )

# Define drug variable lists
drug_vars <- c("CIGFLAG", "CGRFLAG", "PIPFLAG", "SMKLSSFLAG", "NICVAPFLAG", 
               "ALCFLAG", "CBDHMPFLAG", "MRJFLAG", "COCFLAG", "CRKFLAG", 
               "HERFLAG", "HALLUCFLAG", "INHALFLAG", "METHAMFLAG", 
               "PNRANYFLAG", "TRQANYFLAG", "STMANYFLAG", "SEDANYFLAG")

soft_drug_vars <- c("CIGFLAG", "CGRFLAG", "PIPFLAG", "SMKLSSFLAG", 
                    "NICVAPFLAG", "ALCFLAG", "CBDHMPFLAG", "MRJFLAG")

hard_drug_vars <- c("COCFLAG", "CRKFLAG", "HERFLAG", "HALLUCFLAG", 
                    "INHALFLAG", "METHAMFLAG", "PNRANYFLAG", "TRQANYFLAG", 
                    "STMANYFLAG", "SEDANYFLAG")

# Create composite drug use variables
s_clean <- s %>%
  mutate(
    # Main Exposure: Any drug use ever
    any_drug_ever_used = case_when(
      if_any(all_of(drug_vars), ~ . == 1) ~ 1,
      if_all(all_of(drug_vars), ~ . == 0) ~ 0,
      TRUE ~ NA_real_
    ),
    
    # Secondary: Soft drugs
    any_soft_drug_ever_used = case_when(
      if_any(all_of(soft_drug_vars), ~ . == 1) ~ 1,
      if_all(all_of(soft_drug_vars), ~ . == 0) ~ 0,
      TRUE ~ NA_real_
    ),
    
    # Secondary: Hard drugs
    any_hard_drug_ever_used = case_when(
      if_any(all_of(hard_drug_vars), ~ . == 1) ~ 1,
      if_all(all_of(hard_drug_vars), ~ . == 0) ~ 0,
      TRUE ~ NA_real_
    )
  )

# Preview the composite variables
glimpse(s_clean %>% select(any_drug_ever_used, any_soft_drug_ever_used, any_hard_drug_ever_used))

# Define confounders
confounder_vars <- c("AGE3", "IRSEX", "NEWRACE2", "EDUSCHGRD2", "POVERTY3", "AVGGRADE", "PARLMTSN",
                     "EDFAM18", "IMOTHER", "IFATHER", "YEYARGUP", "YEPPROUD",
                      "YEYFGTSW", "YEPGDJOB", "YESCHACT", "YECOMACT")

# Prepare analytic dataset for youth
youth_analytic <- s_clean %>%
  # Filter for youth (ages 12-17)
  filter(AGE3 %in% c(1, 2, 3)) %>%
  
  # Recode variables
  mutate(
    # Recode outcome: Major Depressive Episode (MDE)
    YMDELT_factor = factor(YMDELT, 
                           levels = c(1, 2), 
                           labels = c("Yes", "No")),
    
    # Recode demographics
    AGE3 = factor(AGE3, 
                  levels = c(1, 2, 3), 
                  labels = c("12-13", "14-15", "16-17")),
    
    IRSEX = factor(IRSEX, 
                   levels = c(1, 2), 
                   labels = c("Male", "Female")),
    
    any_drug_ever_used_factor = factor(any_drug_ever_used, 
                                       levels = c(0, 1), 
                                       labels = c("Never Used", "Ever Used"))
  ) %>%
  
  # List-wise deletion: remove NAs in outcome, exposure, and confounders
  drop_na(YMDELT_factor, any_drug_ever_used_factor, all_of(confounder_vars))

# Check sample sizes and distributions (unweighted)
cat("--- Unweighted Age Distribution ---\n")
print(freq(youth_analytic$AGE3))

cat("\n--- Unweighted Drug Use Distribution ---\n")
print(freq(youth_analytic$any_drug_ever_used_factor))

cat("\n--- Unweighted Depression Outcome Distribution ---\n")
print(freq(youth_analytic$YMDELT_factor))



youth_analytic <- youth_analytic %>% 
  mutate(
    EDUSCHGRD2 = ifelse(EDUSCHGRD2>10, NA, EDUSCHGRD2))

youth_analytic <- youth_analytic %>% 
  mutate(
    EDUSCHGRD2 = as.factor(EDUSCHGRD2),
    NEWRACE2   = as.factor(NEWRACE2))

youth_analytic <- youth_analytic %>% 
  mutate(
    depepisode = ifelse(YMDELT_factor=="Yes", 1, 0)
  )
```



```{r survey_design}
# Create survey design object
svy_analytic <- svydesign(
  id      = ~VEREP,
  strata  = ~VESTR_C,
  weights = ~ANALWT2_C,
  data    = youth_analytic,
  nest    = TRUE
)

# Print survey design summary
print(svy_analytic)
```
# unweighted counts & percents
```{r table1}
# Define variable labels
var_labels <- list(
  AGE3 ~ "Age Group",
  IRSEX ~ "Sex",
  NEWRACE2 ~ "Race/Ethnicity",
  EDUSCHGRD2 ~ "Education",
  POVERTY3 ~ "Poverty Level",
  AVGGRADE ~ "Average Grades",
  PARLMTSN ~ "Parental Monitoring")

# Generate unweighted Table 1 stratified by outcome
table1 <- svy_analytic %>%
  tbl_svysummary(
    include = c(AGE3, IRSEX, NEWRACE2, EDUSCHGRD2, POVERTY3, AVGGRADE, 
                PARLMTSN),
    label = var_labels,
    statistic = list(all_categorical() ~ "{n_unweighted} ({p}%)")
  ) %>%
  add_overall(
    col_label = "**Overall (N = {N_unweighted})**",
    statistic = list(all_categorical() ~ "{n_unweighted} ({p}%)")
  ) %>%
  modify_header(label = "**Characteristic**") %>%

# Display the table
table1
```

# Weighted counts, percents, CIs
```{r}

# Weighted counts
weighted_counts <- svytable(~AGE3, svy_analytic)
weighted_counts
# Weighted proportions & 95% CI:
svyprop <- svymean(~factor(AGE3), design = svy_analytic)
svyprop
confint(svyprop)

# Combine results into a table
tbl <- data.frame(
  AGE3 = names(weighted_counts),
  Weighted_Count = as.numeric(weighted_counts),
  Proportion = coef(svyprop),
  SE = SE(svyprop),
  CI_Lower = confint(svyprop)[, 1],
  CI_Upper = confint(svyprop)[, 2]
)

tbl

########################################################################################################

# Weighted counts
weighted_counts <- svytable(~IRSEX, svy_analytic)
weighted_counts
# Weighted proportions & 95% CI:
svyprop <- svymean(~factor(IRSEX), design = svy_analytic)
svyprop
confint(svyprop)


# Combine results into a table
tbl <- data.frame(
  IRSEX = names(weighted_counts),
  Weighted_Count = as.numeric(weighted_counts),
  Proportion = coef(svyprop),
  CI_Lower = confint(svyprop)[, 1],
  CI_Upper = confint(svyprop)[, 2]
)

tbl

########################################################################################################

# Weighted counts
weighted_counts <- svytable(~EDUSCHGRD2, svy_analytic)
weighted_counts
# Weighted proportions & 95% CI:
svyprop <- svymean(~factor(EDUSCHGRD2), design = svy_analytic)
svyprop
confint(svyprop)


# Combine results into a table
tbl <- data.frame(
  EDUSCHGRD2 = names(weighted_counts),
  Weighted_Count = as.numeric(weighted_counts),
  Proportion = coef(svyprop),
  CI_Lower = confint(svyprop)[, 1],
  CI_Upper = confint(svyprop)[, 2]
)

tbl

########################################################################################################

# Weighted counts
weighted_counts <- svytable(~NEWRACE2, svy_analytic)
weighted_counts
# Weighted proportions & 95% CI:
svyprop <- svymean(~factor(NEWRACE2), design = svy_analytic)
svyprop
confint(svyprop)


# Combine results into a table
tbl <- data.frame(
  NEWRACE2 = names(weighted_counts),
  Weighted_Count = as.numeric(weighted_counts),
  Proportion = coef(svyprop),
  CI_Lower = confint(svyprop)[, 1],
  CI_Upper = confint(svyprop)[, 2]
)

tbl

########################################################################################################

# Weighted counts
weighted_counts <- svytable(~POVERTY3, svy_analytic)
weighted_counts
# Weighted proportions & 95% CI:
svyprop <- svymean(~factor(POVERTY3), design = svy_analytic)
svyprop
confint(svyprop)


# Combine results into a table
tbl <- data.frame(
  POVERTY3 = names(weighted_counts),
  Weighted_Count = as.numeric(weighted_counts),
  Proportion = coef(svyprop),
  CI_Lower = confint(svyprop)[, 1],
  CI_Upper = confint(svyprop)[, 2]
)

tbl

########################################################################################################

# Weighted counts
weighted_counts <- svytable(~AVGGRADE, svy_analytic)
weighted_counts
# Weighted proportions & 95% CI:
svyprop <- svymean(~factor(AVGGRADE), design = svy_analytic)
svyprop
confint(svyprop)


# Combine results into a table
tbl <- data.frame(
  AVGGRADE = names(weighted_counts),
  Weighted_Count = as.numeric(weighted_counts),
  Proportion = coef(svyprop),
  CI_Lower = confint(svyprop)[, 1],
  CI_Upper = confint(svyprop)[, 2]
)

tbl

########################################################################################################

# Weighted counts
weighted_counts <- svytable(~PARLMTSN, svy_analytic)
weighted_counts
# Weighted proportions & 95% CI:
svyprop <- svymean(~factor(PARLMTSN), design = svy_analytic)
svyprop
confint(svyprop)


# Combine results into a table
tbl <- data.frame(
  PARLMTSN = names(weighted_counts),
  Weighted_Count = as.numeric(weighted_counts),
  Proportion = coef(svyprop),
  CI_Lower = confint(svyprop)[, 1],
  CI_Upper = confint(svyprop)[, 2]
)

tbl

```

# Confounder check - any drug
```{r confounder_tests}
# Define variables for testing
outcome_var <- "YMDELT_factor"
exposure_var <- "any_drug_ever_used_factor"
confounders_to_test <- c("AGE3", "IRSEX", "NEWRACE2", "EDUSCHGRD2", 
                         "POVERTY3", "AVGGRADE", "PARLMTSN",
                         "EDFAM18", "IMOTHER",    "IFATHER",    "YEYARGUP",
                         "YEPPROUD",   "YEYFGTSW",   "YEPGDJOB",  "YESCHACT", "YECOMACT" )



# Function to safely run svychisq and handle errors
safe_svychisq <- function(formula, design) {
  tryCatch({
    svychisq(formula, design = design)
  }, error = function(e) {
    message("Error in svychisq for ", deparse(formula), ": ", e$message)
    return(NULL)
  })
}

# Test confounders vs. outcome
cat("--- Confounder vs. Outcome (Depression) Tests ---\n")
conf_outcome_tests <- lapply(confounders_to_test, function(conf) {
  formula <- as.formula(paste("~", conf, "+", outcome_var))
  safe_svychisq(formula, svy_analytic)
})
names(conf_outcome_tests) <- confounders_to_test
print(conf_outcome_tests)

# Test confounders vs. exposure
cat("\n\n--- Confounder vs. Exposure (Drug Use) Tests ---\n")
conf_exposure_tests <- lapply(confounders_to_test, function(conf) {
  formula <- as.formula(paste("~", conf, "+", exposure_var))
  safe_svychisq(formula, svy_analytic)
})
names(conf_exposure_tests) <- confounders_to_test
print(conf_exposure_tests)
```
# Regression - Aim 1 - SB version, ORs and CIs not printing
```{r}

# CRUDE
m1 <- svyglm(depepisode ~ as.factor(any_drug_ever_used_factor), design = svy_analytic, family = quasibinomial)
summary(m1)
# print OR and 95% CI
ci1 <- as.table(exp(cbind(OR = coef(m1), confint(m1))))
logLik(m1)


###################

# Fit survey-weighted logistic regression
m1 <- svyglm(depepisode ~ as.factor(any_drug_ever_used_factor),
             design = svy_analytic,
             family = quasibinomial())
 
summary(m1)
 
# ---- Odds Ratios and 95% CIs ----
# Extract coefficients + CI on log-odds scale
logit_ci <- confint(m1)
 
# Convert to OR scale
or_ci <- exp(cbind(OR = coef(m1), logit_ci))
 
# Print nicely
or_ci

########################

# Epoch 1 - age, sex (biological)
m2 <- svyglm(depepisode ~ as.factor(any_drug_ever_used_factor) + AGE3 + IRSEX, design = svy_analytic, family = quasibinomial)
summary(m2)
# print OR and 95% CI
ci2 <- as.table(exp(cbind(OR = coef(m2), confint(m2))))
logLik(m2)

# Epoch 2 - race/ethnicity, education (social)
m3 <- svyglm(depepisode ~ as.factor(any_drug_ever_used_factor) + AGE3 + IRSEX + NEWRACE2 + EDUSCHGRD2, design = svy_analytic, family = quasibinomial)
summary(m3)
# print OR and 95% CI
ci3 <- as.table(exp(cbind(OR = coef(m3), confint(m3))))
logLik(m3)

# Epoch 3 - parent stuff (behavioral)
m4 <- svyglm(depepisode ~ as.factor(any_drug_ever_used_factor) + AGE3 + IRSEX + NEWRACE2 + EDUSCHGRD2 +
               as.factor(YEYARGUP) + as.factor(YEPPROUD) + as.factor(YEYFGTSW) + 
               as.factor(YEPGDJOB), design = svy_analytic, family = quasibinomial)
summary(m4)
# print OR and 95% CI
ci4 <- as.table(exp(cbind(OR = coef(m4), confint(m4))))
logLik(m4)

```

#regression, any drug every used
```{r}
###check to see which confounders stay and which go

########crude model adeu
# 1. Define the Adjusted Logistic Regression Model
m_adeucrude <- svyglm(depepisode ~ as.factor(any_drug_ever_used_factor), 
                      design = svy_analytic, 
                      family = quasibinomial)
# 2. Get the model summary (Log-Odds)
summary(m_adeucrude)
# 3. Extract the Log-Odds Estimates (Coefficients)
log_odds <- coef(m_adeucrude)
# 4. Extract the 95% Confidence Intervals for the Log-Odds
log_ci <- confint(m_adeucrude)
# 5. Combine the estimates and CIs into a single matrix
adeu_log <- cbind(
    Log_Odds = log_odds,
    Lower_CI = log_ci[, 1],
    Upper_CI = log_ci[, 2]
)
# 6. Exponentiate the entire matrix to get the Odds Ratios and their 95% CIs
adeu_OR_CI <- exp(adeu_log)
# 7. Print the final table of Odds Ratios and CIs
kable(adeu_OR_CI)

#########biological model adeu
m_adeubiological <- svyglm(
    formula = depepisode ~ as.factor(any_drug_ever_used_factor) + AGE3 + IRSEX, 
    design = svy_analytic, 
    family = quasibinomial
)
summary(m_adeubiological)
log_odds <- coef(m_adeubiological)
log_ci <- confint(m_adeubiological)
adeubio_log <- cbind(
    Log_Odds = log_odds,
    Lower_CI = log_ci[, 1],
    Upper_CI = log_ci[, 2]
)
adeubio_OR_CI <- exp(adeubio_log)
kable(adeubio_OR_CI)

#########social model adeu
m_adeusocial <- svyglm(depepisode ~ as.factor(any_drug_ever_used_factor), + AGE3 + IRSEX + NEWRACE2 + EDUSCHGRD2 + POVERTY3,
                      design = svy_analytic, 
                      family = quasibinomial,
                      )
summary(m_adeusocial)
log_odds <- coef(m_adeusocial)
log_ci <- confint(m_adeusocial)
adeusoc_log <- cbind(
    Log_Odds = log_odds,
    Lower_CI = log_ci[, 1],
    Upper_CI = log_ci[, 2]
)
adeusoc_OR_CI <- exp(adeusoc_log)
kable(adeusoc_OR_CI)

#########behavioral model adeu
m_adeubeh <- svyglm(depepisode ~ as.factor(any_drug_ever_used_factor), + AGE3 + IRSEX, + NEWRACE2 + EDUSCHGRD2 + POVERTY3 + AVGGRADE + PARLMTSN + EDFAM18 + IMOTHER + IFATHER + YEYARGUP + YEPPROUD +
    YEYFGTSW + YEPGDJOB + YESCHACT + YECOMACT,
                      design = svy_analytic, 
                      family = quasibinomial)
summary(m_adeubeh)
log_odds <- coef(m_adeubeh)
log_ci <- confint(m_adeubeh)
adeubeh_log <- cbind(
    Log_Odds = log_odds,
    Lower_CI = log_ci[, 1],
    Upper_CI = log_ci[, 2]
)
adeubeh_OR_CI <- exp(adeubeh_log)
kable(adeubeh_OR_CI)



```

# regression, soft drug ever used
```{r}

# 1. Define the Adjusted Logistic Regression Model
m_sdeubiological <- svyglm(
    formula = depepisode ~ as.factor(any_soft_drug_ever_used) + AGE3 + IRSEX, 
    design = svy_analytic, 
    family = quasibinomial
)

# 2. Get the model summary (Log-Odds)
summary(m_sdeubiological)

# --- Odds Ratio (OR) and 95% Confidence Interval (CI) Calculation ---

# 3. Extract the Log-Odds Estimates (Coefficients)
log_odds <- coef(m_sdeubiological)

# 4. Extract the 95% Confidence Intervals for the Log-Odds
#    The confint() function correctly accounts for the survey design
log_ci <- confint(m_sdeubiological)

# 5. Combine the estimates and CIs into a single matrix
sdeu_log <- cbind(
    Log_Odds = log_odds,
    Lower_CI = log_ci[, 1],
    Upper_CI = log_ci[, 2]
)

# 6. Exponentiate the entire matrix to get the Odds Ratios and their 95% CIs
#    The exp() function converts Log-Odds to Odds Ratios
sdeu_OR_CI <- exp(sdeu_log)

# 7. Print the final table of Odds Ratios and CIs
print(sdeu_OR_CI)

```

# Flip aim 1
```{r}


# CRUDE
m1 <- svyglm(any_drug_ever_used_factor ~ as.factor(YMDELT_factor), design = svy_analytic, family = quasibinomial)
summary(m1)
# print OR and 95% CI
ci1 <- as.table(exp(cbind(OR = coef(m1), confint(m1))))
logLik(m1)

# Epoch 1 - age, sex (biological)
m2 <- svyglm(any_drug_ever_used_factor ~ as.factor(YMDELT_factor) + AGE3 + IRSEX, design = svy_analytic, family = quasibinomial)
summary(m2)
# print OR and 95% CI
ci2 <- as.table(exp(cbind(OR = coef(m2), confint(m2))))
logLik(m2)



```




```{R}

# Function to fit and summarize weighted logistic regression
fit_logistic <- function(exposure, confounders, design) {
  formula <- as.formula(paste(outcome_var, "~", exposure, "+", paste(confounders, collapse = " + ")))
  model <- svyglm(formula, design = design, family = quasibinomial())
  tidy(model, conf.int = TRUE, exponentiate = TRUE)  # Odds ratios with CIs
}

# Fit models for each exposure
models <- list()
for (exp in all_exposures) {
  cat("\n\n--- Logistic Regression for", exp, "vs. MDE ---\n")
  models[[exp]] <- fit_logistic(exp, confounders_to_test, svy_analytic)
  print(models[[exp]])
}

# Export model results
for (exp in names(models)) {
  write.csv(models[[exp]], paste0("logistic_", exp, "_results.csv"), row.names = FALSE)
}
```


```


```{r debugging, eval=FALSE}
# Debugging: Check for variables with only one level (run if errors occur)
# Uncomment and run this chunk if tbl_svysummary or svychisq fails

cat("Checking categorical variables for single levels:\n")

check_vars <- c("AGE3", "IRSEX", "NEWRACE2", "EDUSCHGRD2", "POVERTY3", 
                "AVGGRADE", "PARLMTSN", "YMDELT_factor", "any_drug_ever_used_factor")

for (var in check_vars) {
  cat("\n---", var, "---\n")
  print(table(youth_analytic[[var]], useNA = "ifany"))
}
```
