---
title: "youth and drug use"
output: html_document
date: "2025-10-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, load packages}

library(tidyverse)
library(summarytools)
library(survey)
library(dplyr)
```

```{r, load data}

load("~/code/NSDUH_2023.Rdata")
library(readr)
```

```{r, renaming}

 d <-"~/code/NSDUH_2023.Rdata"

 vars <-as_tibble(colnames(d))
```

#variables of interest, WHAT AGES DO RANGES ACTUALLY CORRELATE TO?
#demographics/confounders: QUESTID2, AGE3, CATAG2 (recoded 12-17=1), IRSEX, NEWRACE2, EDUHIGHCAT, HEALTH2 ??, IRFAMIN3 OR POVERTY3 ??,
#soft drugs: CIGFLAG, CGRFLAG, PIPFLAG, SMKLSSFLAG, NICVAPFLAG, ALCFLAG, CBDHMPFLAG, MRJFLAG, 0=NEVER USED, 1=EVER USED
#hard drugs: COCFLAG, CRKFLAG, HERFLAG, HALLUCFLAG, INHALFLAG, METHAMFLAG, PNRANYFLAG, TRQANYFLAG, STMANYFLAG, SEDANYFLAG, 0=NEVER USED, 1=EVER USED
#youth experiences aka suicidal thoughts in past year: YUSUITHKYR, .=18+/UNK, 1=YES, 2=NO, 3=IDK, 4=REFUSED
#adolescent depression pg504: lifetime mde YMDELT, .=+18/UNK, 1=YES, 2=NO; past year mde and alc use disorder YMDEAUD5YR, past year mde and illicit drug use disorder YMIUD5YANY, past year mde and substance use disorder YMSUD5YANY, .=+18/UNK, 0=NO, 1=YES

```{r, new data with vars of interest}

 s <- data %>%
   select(c(QUESTID2, AGE3, CATAG2, IRSEX, NEWRACE2, EDUHIGHCAT, POVERTY3, CIGFLAG, CGRFLAG, PIPFLAG, SMKLSSFLAG, NICVAPFLAG, ALCFLAG, CBDHMPFLAG, MRJFLAG, COCFLAG, CRKFLAG, HERFLAG, HALLUCFLAG, INHALFLAG, METHAMFLAG, PNRANYFLAG, TRQANYFLAG, STMANYFLAG, SEDANYFLAG, YMDELT, YMDEAUD5YR, YMIUD5YANY, YMSUD5YANY))
```

```{r, composite variable of all drugs ever used}

drug_vars <- c("CIGFLAG", "CGRFLAG", "PIPFLAG", "SMKLSSFLAG", "NICVAPFLAG", "ALCFLAG", "CBDHMPFLAG", "MRJFLAG", "COCFLAG", "CRKFLAG", "HERFLAG", "HALLUCFLAG", "INHALFLAG", "METHAMFLAG", "PNRANYFLAG", "TRQANYFLAG", "STMANYFLAG", "SEDANYFLAG")

ever_used <- s %>%
  mutate(
    any_drug_ever_used = ifelse(rowSums(select(., all_of(drug_vars))) > 0, 1, 0)
  )

cat("--- Column names in the 'ever_used' dataframe ---\n")
names(ever_used)

```

```{r, composite variable filtered to only have <18}

youth_ever_used <- ever_used %>%
  filter(AGE3 %in% c(1, 2, 3))

cat("\n--- Age distribution in the new 'youth_ever_used' dataframe ---\n")
table(youth_ever_used$AGE3)

cat("\n--- Frequency table for 'any_drug_ever_used' in the youth subset ---\n")
table(youth_ever_used$any_drug_ever_used)

```
